An Efficient Minibatch Acceptance Test for Metropolis-Hastings

### 2016-10-03 nips_2016.tex
Few typos

We run MCMC with our MH test using minibatch size $m=100$.  We also run this
using the methods from~\cite{cutting_mh_2014} (``Korattikara 2014'')
and~\cite{icml2014c1_bardenet14} (``Bardenet 2014''), also with $m=100$.
For~\cite{cutting_mh_2014}, we increment minibatches by 100 as recommended, and
for~\cite{icml2014c1_bardenet14}, we increase the minibatch size geometrically
with a ratio of $\gamma = 1.5$ as recommended. These tests use different
minibatch size updates because of their different approaches to repeated
testing. \cite{icml2014c1_bardenet14} uses straightforward discounting,
allocating a fraction of the test error $\delta$ to each minibatch test (call
these subtests). Its important to bound the number of subtests therefore, so
that the tolerable error per subtest is not too small.  \cite{cutting_mh_2014}
uses a more aggressive repeated testing approach, explicitly modeling
correlation between tests and allowing many more tests for a given $\delta$.

The tolerance for making a decision in~\cite{cutting_mh_2014} is
$\epsilon=0.005$, and the error bound control parameters
in~\cite{icml2014c1_bardenet14} are $p = 2$ and $\delta = 0.01$.  All methods
(including ours) use the same random walk proposer with covariance $\Sigma={\rm
diag}(0.3, 0.3)$.  Since the proposer generates a random walk, all shaping of
distribution of the samples is due to the MH test. All methods are run 5000
times to collect 5000 samples.

@@

In practice, it is difficult to compute multinomial probabilities for large $n$,
in that case the per-bin distributions are well-approximated by Poisson
distributions with parameter $\lambda_i=P_i n$. Given a set of samples $\Theta =
\{\theta_1,\ldots,\theta_T\}$, let $c_j$ denote the number of individual samples
$\theta_i$ that fall in bin $j$. Then the log probability for the set is
approximated as
\begin{equation}\label{eq:log_prob_poissons}
    \log p(\Theta) \approx \frac{1}{N_{\rm bins}} \sum_{j=1}^{N_{\rm bins}} c_j \log (n P_j) - n P_j - \log(\Gamma(c_j+1)).
\end{equation}
The results for each model are shown in Table~\ref{tab:poissons} ...
{\color{blue} Daniel: IN PROGRESS}

While the Poisson results provide general guidance on the relative accuracy of
the models it is difficult interpret the quantitative scores. We therefore
perform significance tests to test the null hypothesis that the ideal
distribution really generated the samples.  We perform two tests:

\begin{enumerate}
    \item An exact test using simulation to generate an approximate distribution
    of scores given the null hypothesis.

    \item An approximate test using the chi-squared distribution as the test
    statistic. 
\end{enumerate}

The results were... {\color{blue} Daniel: IN PROGRESS}

@@

\begin{table}[t]
    \caption{\color{blue} Daniel: TODO Poisson Probabilities}
    \label{tab:poissons}
    \centering
    \begin{tabular}{l l l l}
    \toprule
    Model & Ours & Korattikara 2014 & Bardenet 2014 \\
    \midrule
    Poisson Value (Eq.~\ref{eq:log_prob_poissons}) & TODO & TODO & TODO \\
    \bottomrule
    \end{tabular}
\end{table}

@@

@@

>>>>

We run MCMC with our MH test using minibatch size $m=100$.  We also run this
using the methods from~\cite{cutting_mh_2014} (``Korattikara 2014'')
and~\cite{icml2014c1_bardenet14} (``Bardenet 2014''), also with $m=100$.
For~\cite{cutting_mh_2014}, we increment minibatches by 100 as recommended, and
for~\cite{icml2014c1_bardenet14}, we increase the minibatch size geometrically
with a ratio of $\gamma = 1.5$ as recommended. These tests use different
minibatch size updates because of their different approaches to repeated
testing. \cite{icml2014c1_bardenet14} uses straightforward discounting,
allocating a fraction of the test error $\delta$ to each minibatch test (call
these subtests). It's important to bound the number of subtests therefore, so
that the tolerable error per subtest is not too small.  \cite{cutting_mh_2014}
uses a more aggressive repeated testing approach, explicitly modeling
correlation between tests and allowing many more tests for a given $\delta$.

The tolerance for making a decision in~\cite{cutting_mh_2014} is
$\epsilon=0.005$, and the error bound control parameters
in~\cite{icml2014c1_bardenet14} are $p = 2$ and $\delta = 0.01$.  All methods
(including ours) use the same random walk proposer with covariance $\Sigma={\rm
diag}(0.15, 0.15)$.  Since the proposer generates a random walk, all shaping of
distribution of the samples is due to the MH test. All methods are run 5000
times to collect 5000 samples.

@@

In practice, it is difficult to compute multinomial probabilities for large $n$,
in that case the per-bin distributions are well-approximated by Poisson
distributions with parameter $\lambda_i=P_i n$. Given a set of samples $\Theta =
\{\theta_1,\ldots,\theta_T\}$, let $c_j$ denote the number of individual samples
$\theta_i$ that fall in bin $j$ out of $N_{\rm bins}$ total. We omit bins in
which either the source or MCMC-generated distributions did not have a single
sample, resulting in the set of (at most $N_{\rm bins}$) bins $B$. Then the log
probability for the set of samples is approximated as
\begin{equation}\label{eq:log_prob_poissons}
    %\log p(\Theta) \approx \frac{1}{N_{\rm bins}} \sum_{j=1}^{N_{\rm bins}} c_j \log (n P_j) - n P_j - \log(\Gamma(c_j+1)).
    \log p(\Theta) \approx \sum_{j \in B} c_j \log (n P_j) - n P_j - \log(\Gamma(c_j+1)).
\end{equation}
The results for each model are shown in Table~\ref{tab:poissons}.

While the Poisson results provide general guidance on the relative accuracy of
the models, it is difficult to interpret the quantitative scores. We therefore
perform significance tests to show the difference between the MCMC-sampled
distributions and the ground-truth.  We perform an approximate test using
the Chi-Squared distribution as the test statistic, which we also show in
Table~\ref{tab:poissons}.

Both the Poisson values and the Chi-Squared test statistic result in similar
conclusions: our method is slightly superior to~\cite{cutting_mh_2014}, but
slightly worse than~\cite{icml2014c1_bardenet14}. We now show, however, that
minibatch size results suggest that our method dominates in terms of speed and
efficiency.

@@

\begin{table}[t]
    \caption{Gaussian Mixture Model Statistics}
    \label{tab:poissons}
    \centering
    \begin{tabular}{l l l l}
    \toprule
    Model & Our Method & Korattikara 2014 & Bardenet 2014 \\
    \midrule
    Poisson Value (Eq.~\ref{eq:log_prob_poissons}) & -1430.0 & -1578.9 & -1232.7 \\
    Chi-Squared Test Statistic & 3313.9 & 3647.7 & 2444.1 \\
    \bottomrule
    \end{tabular}
\end{table}

@@

\begin{comment}

@@

\end{comment}
